---
title: "Biological Network Analysis - class 17"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Practice data
```{r}
cwd <- demoSimpleGraph()

layoutNetwork( cwd, 'force-directed')

possible.layout.names <- getLayoutNames(cwd)
layoutNetwork (cwd, possible.layout.names[1])

ping(cwd)
```

Interact with cytoscape to change visuals - TEST

```{r}
setVisualStyle(cwd, "Marquee")

styles <- getVisualStyleNames(cwd)
styles

setVisualStyle(cwd, styles[13])
setVisualStyle(cwd, styles[11])
```

Read in .tsv file
```{r}
prok_vir_cor <- read.delim("virus_prok_cor_abundant.tsv", stringsAsFactors = FALSE)

head(prok_vir_cor)

```

Using igraph, create a network. directed = false because we don't know the directionality of the interactions
```{r}
g <- graph.data.frame(prok_vir_cor, directed = FALSE)
class(g)
g
```

Plot the data
```{r}
plot(g)
```

For something a little more useful
```{r}
plot(g, vertex.size=3, vertex.label=NA)
```

To list or alter details about the edges and vertices:
```{r}
V(g)
E(g)
```

Read in some more data
```{r}
phage_id_affiliation <- read.delim("phage_ids_with_affiliation.tsv")
head(phage_id_affiliation)

bac_id_affi <- read.delim("prok_tax_from_silva.tsv")
head(bac_id_affi)
```

Now we can create a network
```{r}
genenet.nodes <- as.data.frame(vertex.attributes(g))

## not all have classification, so create empty columns
genenet.nodes$phage_aff <- rep("not_class", nrow(genenet.nodes))
genenet.nodes$Tax_order <- rep("not_class", nrow(genenet.nodes))
genenet.nodes$Tax_subfamily <- rep("not_class", nrow(genenet.nodes))

for (row in seq_along(1:nrow(genenet.nodes))){
  if (genenet.nodes$name[row] %in% phage_id_affiliation$first_sheet.Phage_id_network){
    id_name <- as.character(genenet.nodes$name[row])
    aff_to_add <- unique(subset(phage_id_affiliation,
                                first_sheet.Phage_id_network == id_name,
                                select = c(phage_affiliation,
                                           Tax_order,
                                           Tax_subfamily)))
    genenet.nodes$phage_aff[row] <- as.character(aff_to_add$phage_affiliation)
    genenet.nodes$Tax_order[row] <- as.character(aff_to_add$Tax_order)
    genenet.nodes$Tax_subfamily[row] <- as.character(aff_to_add$Tax_subfamily)
  }
}

## do the same for proks
genenet.nodes$prok_king <- rep("not_class", nrow(genenet.nodes))
genenet.nodes$prok_tax_phylum <- rep("not_class", nrow(genenet.nodes))
genenet.nodes$prok_tax_class <- rep("not_class", nrow(genenet.nodes))

for (row in seq_along(1:nrow(genenet.nodes))){
  if (genenet.nodes$name[row] %in% bac_id_affi$Accession_ID){
    aff_to_add <- unique(subset(bac_id_affi,
                                Accession_ID == as.character(genenet.nodes$name[row]),
                                select = c(Kingdom,
                                           Phylum,
                                           Class)))
    
    genenet.nodes$prok_king[row] <- as.character(aff_to_add$Kingdom)
    genenet.nodes$prok_tax_phylum[row] <- as.character(aff_to_add$Phylum)
    genenet.nodes$prok_tax_class[row] <- as.character(aff_to_add$Class)
  }
}

genenet.edges <- data.frame(igraph::as_edgelist(g))
names(genenet.edges) <- c("name.1", "name.2")

genenet.edges$Weight <- igraph::edge_attr(g)[[1]]

genenet.edges$name.1 <- as.character(genenet.edges$name.1)
genenet.edges$name.2 <- as.character(genenet.edges$name.2)
genenet.nodes$name <- as.character(genenet.nodes$name)

ug <- cyPlot(genenet.nodes,genenet.edges)
```

Send to cytoscape using RCy3
```{r}
cy <- CytoscapeConnection()
deleteAllWindows(cy)

cw <- CytoscapeWindow("Tara oceans",
                      graph = ug,
                      overwriteWindow = TRUE)

displayGraph(cw)
layoutNetwork(cw)
fitContent(cw)
```

The Network in cytoscape can be manipulated from R
```{r}
families_to_colour <- unique(genenet.nodes$prok_tax_phylum)
families_to_colour <- families_to_colour[!families_to_colour %in% "not_class"]
node.colour <- RColorBrewer::brewer.pal(length(families_to_colour), "Set3")

setNodeColorRule(cw,
                 "prok_tax_phylum",
                 families_to_colour,
                 node.colour,
                 "lookup",
                 default.color = "#ffffff")

displayGraph(cw)
layoutNetwork(cw)
fitContent(cw)
```

Set node shapes
```{r}
shapes_for_nodes <- c("DIAMOND")

phage_names <- grep("ph_",
                    genenet.nodes$name,
                    value = TRUE)
setNodeShapeRule(cw,
                 "label",
                 phage_names,
                 shapes_for_nodes)

displayGraph(cw)
fitContent(cw)
```

Change edges
```{r}
setDefaultNodeBorderWidth(cw, 5)
families_to_colour <- c(" Podoviridae",
                        " Siphoviridae",
                        " Myoviridae")
node.colour <- RColorBrewer::brewer.pal(length(families_to_colour),
                          "Dark2")
setNodeBorderColorRule(cw,
                       "Tax_subfamily",
                       families_to_colour,
                       node.colour,
                       "lookup", 
                       default.color = "#000000")

displayGraph(cw)
fitContent(cw)
```

Layout options
```{r}
getLayoutNames(cw)
getLayoutPropertyNames(cw, layout.name="force-directed")
getLayoutPropertyValue(cw, "force-directed", "defaultSpringLength")
getLayoutPropertyValue(cw, "force-directed", "numIterations")
setLayoutProperties(cw,
                   layout.name = "force-directed",
                    list(defaultSpringLength = 20,
                        "numIterations" = 200))
layoutNetwork(cw,
             layout.name = "force-directed")
fitContent(cw)
layoutNetwork(cw, layout.name = "force-directed")
fitContent(cw)
```

```{r}
## initiate a new node attribute
ug2 <- initNodeAttribute(graph = ug,
                          "degree",
                          "numeric",
                          0.0) 

## degree from graph package for undirected graphs not working well,
## so instead using igraph to calculate this from the original graph
nodeData(ug2, nodes(ug2), "degree") <- igraph::degree(g)

cw2 <- CytoscapeWindow("Tara oceans with degree",
                      graph = ug2,
                      overwriteWindow = TRUE)

displayGraph(cw2)
layoutNetwork(cw2)
```

```{r}
degree_control_points <- c(min(igraph::degree(g)),
                           mean(igraph::degree(g)),
                           max(igraph::degree(g)))
node_sizes <- c(20,
                20,
                80,
                100,
                110) # number of control points in interpolation mode,
                     # the first and the last are for sizes "below" and "above" the attribute seen.

setNodeSizeRule(cw2,
                "degree",
                degree_control_points,
                node_sizes,
                mode = "interpolate")
layoutNetwork(cw2, "force-directed")

displayGraph(cw2)
layoutNetwork(cw2)
```

Select specific nodes
```{r}
selectNodes(cw2, "GQ377772") 
getSelectedNodes(cw2)

selectFirstNeighborsOfSelectedNodes(cw2)
getSelectedNodes(cw2)

selectFirstNeighborsOfSelectedNodes(cw2)
getSelectedNodes(cw2)
```

Create a new network from this selection
```{r}
newnet <- createWindowFromSelection(cw2,
                                    "subnet",
                                    "TRUE")

layoutNetwork(newnet, "force-directed")
```

